Architectures for workflow integration in science gateways

### 2016-04-12 paper.tex
Wrote section 4.2

\todo{Introduce the notion of architectural diagram with an abstract figure. Mention arrows, boxes, box with dotted lines, red color.}

@@

See Table~\ref{table:evaluation}.

We focus on the integration of workflow engines in science gateways,
i.e. we assume that the science gateway is already interfaced with the
infrastructure.

@@

\emph{Scalability} measures the ability of the architecture to support
high workloads. It breaks down into 4 metrics:
\begin{itemize}[itemsep=0cm]
\item New engine instance (0: automated; 1: manual intervention needed).
\item Elastic engines (0: easy to implement; 1: difficult to
  implement; 2: not available).
\item Distributed engines (0: possible; 1: not possible).
\item Task scheduling (0: difficult; 1: more difficult).
\end{itemize}

@@

@@

\begin{table*}
\centering
\begin{tabular}{rcccccc}
\textbf{Metric}                      & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Sub-task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Import} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Integration}}\\
Science gateway -- \texttt{I$_1$}    & \multirow{2}{*}{3}
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}0  
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99AA99}3 \\
Workflow engine -- \texttt{I$_2$}    &
                                     & \cellcolor[HTML]{99FF99}3 
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}1 \\
Infrastructure -- \texttt{I$_3$}   & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
Other -- \texttt{I$_4$}              & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}3
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{4}
                                     & \cellcolor[HTML]{99FF99}\textbf{2}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{4}
                                     & \cellcolor[HTML]{99AA99}\textbf{4} \\

@@

\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{4}

@@

New infrastructure -- \texttt{M$_3$} & \cellcolor[HTML]{99FF99}2

@@

\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{8}

@@

New engine instance -- \texttt{S$_1$}& \cellcolor[HTML]{99AA99}1

@@

\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{4}

@@

                                     & \cellcolor[HTML]{99CC99}\textbf{19}

>>>>

\todo{Introduce the notion of architectural diagram with an abstract
  figure. Mention arrows, boxes, box with dotted lines, red color,
  tasks, data.}

@@

See Table~\ref{table:evaluation}. We focus on the integration of
workflow engines in science gateways, i.e. we assume that the science
gateway is already interfaced with the infrastructure.

@@

\emph{Scalability} measures the ability of the architecture to support
high workloads. It breaks down into 4 metrics:
\begin{itemize}[itemsep=0cm]
\item New engine instance (0: automated; 1: manual intervention needed; 2: not possible).
\item Elastic engines (0: easy to implement; 1: difficult to
  implement; 2: not available).
\item Distributed engines (0: possible; 1: not possible).
\item Task scheduling (0: difficult; 1: more difficult).
\end{itemize}

@@

\paragraph{Integration} Tightly integrating a workflow engine in a
science gateway requires to modify the science gateway
(\texttt{I$_1$=1}) and to develop interactions \texttt{b}, and
\texttt{c$_2$} in the workflow engine (\texttt{I$_2$=2}). No other
modification is required (\texttt{I$_3$}=\texttt{I$_4$}=0).

\paragraph{Robustness} The science gateway and the infrastructure are
involved in a workflow execution (\texttt{R$_1$}=2), with interactions
\texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{R$_2$}=3). Debugging is not an issue since the science
gateway can retrieve any information from the workflow engine
directly (\texttt{R$_3$}=0).

\paragraph{Modularity} Integrating a new type of workflow engine
requires to modify the science gateway as well as interactions
\texttt{b} and \texttt{c$_2$} (\texttt{M$_1$}=3). Updating a workflow
engine version requires modifications in the science gateway
(\texttt{M$_2$}=1). Adding a new infrastructure generates updates in
interactions \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{M$_3$}=3). Inserting a new workflow is done through
interaction \texttt{a} (\texttt{M$_4$}=1). Meta-workflows are not
supported by default (\texttt{M$_4$}=1).

\paragraph{Scalability} Adding a new engine instance requires a new
instance of the science gateway, which is in general not possible
since load-balancing mechanisms available in web servers are usually
meant to deal with short web requests rather than long workflow
executions (\texttt{S$_1$}=2). Coherently, elastic engines are usually
not possible either (\texttt{S$_2$}=2).  Distributed engines are not
available by default (\texttt{S$_3$}=1). The scheduling of tasks on
the infrastructure is as complex as in any other architecture since
the workflow engine might implement any kind of scheduling policy
(\texttt{S$_4$}=0).

@@

\begin{table*}
\centering
\begin{tabular}{rcccccc}
\textbf{Metric}                      & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Sub-task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Import} \\
\multicolumn{7}{c}{\cellcolor[HTML]{EEEEEE}\textbf{Integration}}\\
Science gateway -- \texttt{I$_1$}    & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}0  
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99AA99}3 \\
Workflow engine -- \texttt{I$_2$}    & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}3 
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99AA99}2
                                     & \cellcolor[HTML]{99FF99}1 \\
Infrastructure -- \texttt{I$_3$}   & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
Other -- \texttt{I$_4$}              & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99AA99}3
                                     & \cellcolor[HTML]{99AA99}0
                                     & \cellcolor[HTML]{99FF99}0 \\
\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{3}
                                     & \cellcolor[HTML]{99FF99}\textbf{4}
                                     & \cellcolor[HTML]{99FF99}\textbf{2}
                                     & \cellcolor[HTML]{99AA99}\textbf{6}
                                     & \cellcolor[HTML]{99AA99}\textbf{4}
                                     & \cellcolor[HTML]{99AA99}\textbf{4} \\

@@

\textbf{Total}                       & \cellcolor[HTML]{99FF99}\textbf{5}

@@

New infrastructure -- \texttt{M$_3$} & \cellcolor[HTML]{99FF99}3

@@

\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{9}

@@

New engine instance -- \texttt{S$_1$}& \cellcolor[HTML]{99AA99}2

@@

\textbf{Total}                       & \cellcolor[HTML]{99AA99}\textbf{5}

@@

                                     & \cellcolor[HTML]{99CC99}\textbf{22}
